---
title: "Data description"
author: "Keisuke Atsumi"
date: '`r format(Sys.time(), "%y/%m/%d %H:%M")`'
output:
  html_document:
    code_folding: hide
    theme: united
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}

rm(list=ls())  # reset workspace
options(scipen=100)  # do not show numbers using exponential

# install & load packages
# install.packages("tidyverse" ,"metafor", "openxlsx", "data.table")
library(tidyverse)  # dataset modification
library(knitr)
library(svglite)
opts_chunk$set(prompt=TRUE, message=FALSE, comment="", warning = FALSE) 

mytheme <- 
  theme(
    panel.grid = element_blank(),
    axis.title = element_text(size=12),
    axis.text = element_text(size=11),
    axis.text.x = element_text(angle = 30, hjust = 1),
    strip.text = element_text(size=12, lineheight=5.0),
    strip.background = element_rect(fill = "lightgrey")
  )


```

# Summarizing dataset
Summarize data for variation only because the data will be the same for mean data for following reasons:   
- variation data removes NAs for variance measures which did not yet for mean data, though NAs for variance measures will be elimated later in mean data as well    
- no data was removed in variation data other than the reasons above  
*original* : no observations were removed  
*fullcross* : observations with both reciprocal cross ... used in comparing paternal/maternal effect in the analysis *general patterns*  
*fullmodifiers* : observations with all modifiers ... used in transgressive segregation meta-regression  
*full* : observations with both reciprocal cross and all modifiers ... used in meta-regression for paternal/maternal effect in the analysis *general patterns*

```{r mean data}


### original data ###
original <- read.csv("../data/dat.CV.csv", head = TRUE) %>%
  within(
  # Order taxon
  taxa <- ordered(
    taxa,
    levels = c(
      "Teleost", "Amphibian", "Bird", "Mammal",
      "Diptera", "Lepidoptera", "Neuroptera", 
      "Orthoptera", "Coleoptera" 
      )
    )
  ) %>%
  group_by(taxa)

### Observations with both reciprocal crosses ###
fullcross <- original %>%
  filter(reciprocal == "Viable")

### Observations with all modifiers ###
fullmodifiers <- original %>%
  drop_na("divergence.COI", "hetero.sex")
  
### Observations with with both reciprocal crosses & all modifiers ###
full <- original %>%
  filter(reciprocal == "Viable") %>%
  drop_na("divergence.COI", "hetero.sex")
  

for (i in c("original", "fullcross", "fullmodifiers", "full")) {

  summary <- get(i) %>%
    group_by(taxa) %>%
    summarise(
      species.pair = length(unique(species.pair)),
      observations = length(unique(ES.ID)),
      study = length(unique(Study.ID))
      )
  
  print(i)
  print(
    bind_rows(
      summary,
      summary %>%
        summarise_if(is.numeric, sum) %>%
        mutate(taxa = "total")
      ) %>%
      column_to_rownames(var = "taxa")
    )
  
  # Order by taxon
  
  sum.forplot <- summary %>% 
    gather(key = metrics, value = N, -taxa) %>%
    within(
      metrics <- ordered(
        metrics,
        levels = c(
          "study", "species.pair", "observations"
          )
        )
      )
  
  
  plot <- ggplot(sum.forplot, aes(x = "", y = N, fill = taxa)) +
    geom_bar(width = 1, stat = "identity") +
    scale_fill_brewer() +
    facet_grid(scale = "free", rows = "metrics") +
    xlab(i) +
    mytheme
  print(plot)
  
  if(i == "fullmodifiers") {
    ggsave(
      plot = plot, 
      file = "../Analysis/data.fullmodifiers.svg",
      height = 4.5, width = 3.5
    )
  }else{
    }

}


```
