---
title: "QTLs and variability"
author: "Keisuke Atsumi"
date: '`r format(Sys.time(), "%y/%m/%d %H:%M")`'
output:
  html_document:
    code_download: true
    code_folding: hide
    theme: united
    toc: yes
    toc_float: yes
---
## Setup

### Load packages

```{r}

rm(list=ls())  # reset workspace
options(scipen=100)  # do not show numbers using exponential

# install & load packages
pacman::p_load(
  openxlsx   # open excel
  , tidyverse
  , magrittr   # extended pipe
  , metafor    # meta-analysis
  , MCMCglmm   # comparative analysis
  , orchaRd    # Orchard plot - meta-analysis
  , rotl       # open tree of life
  , ape        # Phylogeny
  , phytools   # Phylogeny
  , knitr  
  , kableExtra # nice tables
  , pander     # nice tables
  , svglite    # Export SVG plots
  , ggiraphExtra # Geom predict
  # , R.rsp
)

# Rmarkdown settings
knitr::opts_chunk$set(
  prompt  = FALSE,  # Do not add > or + in inline-code
  message = FALSE, 
  comment = "", 
  warning = FALSE,  # Mute warnings
  tidy    = TRUE
  ) 
options(knitr.kable.NA = '') # Hide NAs in kable table
```


# Dataset

I searched QTL mapping studies through forward citations of primary studies. I could find QTLs only for 10 traits. <br>
Among those, all species pairs were male heterogametic, and both recciprocal hybrids were viable. Genetic distance metrics was not available for the crossin using *Laupara* crickets.

```{r, results = 'asis'}

## Phenotype data ##
pheno <- read.xlsx(
    "../data/original.data.xlsx", sheet = "Phenotype"
    # Same data is also saved as "../data/original.data.Phenotype.txt"
    ) %>%
  # Excluding data without any parent data
  drop_na(contains("SD.sp")) %>%
  # to numeric
  mutate_at(
    vars(contains("Mn") | contains("SD") | N.sp1:N.sp2), 
    as.numeric
    ) %>%
  # delete observation with negative mean trait value
  filter(Mn.sp1 > 0 & Mn.sp2 > 0 & homologous == "Yes") %>%   
  # calculate phenotypic difference between parental species
  mutate(
    Pheno.divergence = abs(log(Mn.sp1/Mn.sp2))
    ) %>%
  mutate_at(vars(contains("pheno.div")), scale) %>%
  # describe whether full cross or not
  mutate(
    Reciprocal = ifelse(
      is.na(Mn.hyb12 & Mn.hyb21), 
      "Inviable", "Viable"
      )
    )

meta <- read.xlsx(
    "../data/original.data.xlsx", sheet = "Species.level.moderators"
    # Same data is also saved as "../data/original.data.Species.level.moderators.txt"
    ) %>%
  mutate_at(vars(contains("divergence")), as.numeric) %>%
  # Logalize genetic divergence
  mutate_at(vars(contains("divergence")), log10)


#++++++++++++++++++++++++++++++++++++++++++++++
# Combine phenotypic data and moderators data
#++++++++++++++++++++++++++++++++++++++++++++++
original.dat <- left_join(read.csv("QTL.csv"), pheno) %>%
  drop_na(N.QTL) %>%
  left_join(., meta)
# add unique ID for each effect size (row)
original.dat$ES.ID <- paste("ES", sprintf("%03d",c(1:dim(original.dat)[1])), sep="")


summary <- original.dat %>%
  group_by(taxa) %>%
  summarise(
    'Species pair' = length(unique(species.pair)),
    'Obser- vation' = length(unique(ES.ID)),
    Study = length(unique(Study.ID))
    )


### Table ###
original.dat %>%
  kable("html", digits = 3) %>%
  kable_styling("striped", position = "left") %>%
  scroll_box(width = "80%", height = "300px")

### Plot ###
ggplot(
  # Make tidy summary data
  summary %>%
  gather(key = metrics, value = N, -c(taxa)) %>%
  # Order metrics
  within(
    metrics <- ordered(
      metrics,
      levels = c("Study", "Species pair", "Obser- vation")
      )
    ), 
  aes(x = "", y = N, fill = taxa)) +
  geom_bar(width = 1, stat = "identity") +
  scale_fill_brewer(palette = "YlGn") +
  facet_wrap(
    scale = "free", "metrics",
    # strip text into two lines
    labeller = label_wrap_gen(width = 8)
    ) +
  xlab("")


```


# Meta-regression for phenotypic variation

We quantified relative phenotypic variability of hybrids and the parent with large variability (<i>spLL<sub>V</sub></i>) to the parent with small variability (<i>spSS<sub>V</sub></i>), by using [lnCVR](https://besjournals.onlinelibrary.wiley.com/doi/full/10.1111/2041-210X.12309). We controlled mean-variance relationship by using Coefficient of Variation (CV) as the proxy of phenotypic variability. <br>

We then regressed relative phenotypic variability of hybrids (arrows 2&3 in Fig.1b of [Atsumi et al](https://ecoevorxiv.org/kt3ud/)) by phenotypic variability of parent (arrow 1 - to controll the additive inheritance for phenotypic variability, `spL.variability`) and estimated number of QTLs for the focal traits (`N.QTL`).

```{r}

#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++#
# Align parental species by phenotypic variability size
#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++#

#+++++++++++++++++++++++++++++++++++++++++++++++++#
# sp1 >= sp2 in phenotypic variation (CV)
#+++++++++++++++++++++++++++++++++++++++++++++++++#
# extract
sp1sp2.var <- original.dat %>%
  filter(SD.sp1/Mn.sp1 >= SD.sp2/Mn.sp2) # filtering according to CV of parentals
# change column name from sp1 or sp2 -> spL or spS
names(sp1sp2.var) <- gsub("1", "L", names(sp1sp2.var))
names(sp1sp2.var) <- gsub("2", "S", names(sp1sp2.var))

#+++++++++++++++++++++++++++++++++++++++++++++++++#
# sp1 < sp2 in phenotypic variation (CV)
#+++++++++++++++++++++++++++++++++++++++++++++++++#
# extract
sp2sp1.var <- original.dat %>%
  filter(SD.sp1/Mn.sp1 < SD.sp2/Mn.sp2)
# change column name from sp1 or sp2 -> spS or spL
names(sp2sp1.var) <- gsub("1", "S", names(sp2sp1.var))
names(sp2sp1.var) <- gsub("2", "L", names(sp2sp1.var))



#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++#
# Effect size calculation | lnCVR
#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++#

dat <- bind_rows(sp1sp2.var, sp2sp1.var)

# Order by taxon
dat$taxa <- ordered(
  dat$taxa,
  levels = c("Neuroptera", "Coleoptera", "Diptera", "Lepidoptera", "Orthoptera", "Aves", "Rodentia", "Anura", "Cichliformes")
  )

for (i in c("hybLS", "hybSL", "spL")) { # identify type of hybrids
  for (h in c("spS")) { # identify type of parentals
    assign(
      paste(i, h, sep = "_"),
        escalc(
          measure = "CVR",
          # N of hybrids
          n1i = dat[, paste("N", i, sep = ".")],
          # N of parentals
          n2i = dat[, paste("N", h, sep = ".")],
          # Mean of hybrids
          m1i = dat[, paste("Mn", i, sep = ".")],
          # Mean of parentals
          m2i = dat[, paste("Mn", h, sep = ".")],
          # SD of hybrids
          sd1i = dat[, paste("SD", i, sep = ".")],
          # SD of parentals
          sd2i = dat[, paste("SD", h, sep = ".")],
          data = dat
          ) %>%
          mutate(cross = i) 
    )
  }
}

# Combine all ES datasets
CV.dif <- bind_rows(
  hybLS_spS, hybSL_spS
  ) %>%
  left_join(
    spL_spS %>%
      dplyr::select("ES.ID", "yi") %>%
      rename(spL.variability = yi)
  ) %>%
  drop_na(yi) %>%
  arrange(ES.ID) %>%
  left_join(., dat) 


# Regress by lnRR of phenotypic difference between parentals
model <- rma.mv(
  yi = yi, V = vi,
  data = CV.dif, method = "REML",
  random = list(~1 | spL.name, ~1 | ES.ID),
  mods = ~ N.QTL + spL.variability
  )

### Output result ###
reg <- tibble(
    'Fixed_effects' = row.names(model$beta),
    'Estimate' = model$beta, # regression: estimate
    'LowerCI' = model$ci.lb, # regression: 95ci, 
    'UpperCI'  = model$ci.ub, # regression: 95ci
    'P'        = model$pval
    )  %>%
    as.data.frame() %>%
    mutate(significance = ifelse(P < 0.05, "*", "")) %>%
  mutate_at("Fixed_effects", as.factor)

ggplot(reg, aes(y = Estimate, x = Fixed_effects)) +
  geom_errorbar(
    aes(ymin = LowerCI, ymax = UpperCI, colour=significance), 
    width=.0001
    ) +
  # Color of plots and errorbars
  scale_colour_manual(values = c("grey60", "black")) +
  geom_point(size = 1.5, aes(colour = significance)) +
  scale_fill_manual(values = c("grey60", "black")) +
  scale_x_discrete(limits = rev(reg$Fixed_effects))+ # reverse x-axis
  geom_hline(yintercept = 0, linetype="dashed", size = 0.3) +
  coord_flip() +
  theme(
    panel.grid = element_blank(),
    legend.position = "none",
    panel.border=element_blank(),
    axis.text.x = element_text(size = 11),
    axis.text.y = element_text(size = 12),
    axis.line=element_line(colour = "grey70")
    )

```

The result implies negative relationship between F1 phenotypic variability and QTL counts. <br>

```{r}

# Regress by lnRR of phenotypic difference between parentals
model <- rma.mv(
  yi = yi, V = vi,
  data = CV.dif, method = "REML",
  random = list(~1 | spL.name, ~1 | ES.ID),
  mods = ~ N.QTL + Genet.divergence + spL.variability
  )

### Output result ###
reg <- tibble(
    'Fixed_effects' = row.names(model$beta),
    'Estimate' = model$beta, # regression: estimate
    'LowerCI' = model$ci.lb, # regression: 95ci, 
    'UpperCI'  = model$ci.ub, # regression: 95ci
    'P'        = model$pval
    )  %>%
    as.data.frame() %>%
    mutate(significance = ifelse(P < 0.05, "*", "")) %>%
  mutate_at("Fixed_effects", as.factor)

ggplot(reg, aes(y = Estimate, x = Fixed_effects)) +
  geom_errorbar(
    aes(ymin = LowerCI, ymax = UpperCI, colour=significance), 
    width=.0001
    ) +
  # Color of plots and errorbars
  scale_colour_manual(values = c("grey60", "black")) +
  geom_point(size = 1.5, aes(colour = significance)) +
  scale_fill_manual(values = c("grey60", "black")) +
  scale_x_discrete(limits = rev(reg$Fixed_effects))+ # reverse x-axis
  geom_hline(yintercept = 0, linetype="dashed", size = 0.3) +
  coord_flip() +
  theme(
    panel.grid = element_blank(),
    legend.position = "none",
    panel.border=element_blank(),
    axis.text.x = element_text(size = 11),
    axis.text.y = element_text(size = 12),
    axis.line=element_line(colour = "grey70")
    )

```
After genetic divergence between parents was included (reduce 2 data points), sign of the effect of QTL counts remained negative, but became weaker.
