---
title: "General pattern of phenotypic variation"
author: "Keisuke Atsumi"
date: '`r format(Sys.time(), "%y/%m/%d %H:%M")`'
output:
  html_document:
    code_folding: hide
    theme: united
    toc: yes
    toc_float: yes
---

# Setup 

## Loading packages
```{r setup, include=FALSE}

rm(list=ls())  # reset workspace
sessionInfo()  # show R version etc
options(scipen=100)  # do not show numbers using exponential


# load packages
pacman::p_load(
  tidyverse,  
  metafor,    # Meta-analysis
  knitr,      # rmarkdown
  magrittr,   # Extend pipe
  ape,        # Phylogeny
  phytools,   # Phylogeny
  svglite,    # Export SVG plots
  ggbeeswarm, # bee-swarm plots
  plotly      # interactive plots using ggplot2
)

opts_chunk$set(prompt=TRUE, message=FALSE, comment="", warning = FALSE)

mytheme <- 
  theme(
    panel.grid = element_blank(),
    axis.ticks.y=element_blank(),
    axis.title = element_text(size=20),
    axis.text = element_text(size=17, color = "black"),
    axis.line.y = element_blank(),
    strip.text = element_text(size=12, lineheight=5.0),
    strip.background = element_rect(fill = "white")
  )

```


## Custom functions

```{r custom functions}

# General modeling functions Functions for I2

#' Title Function to obtain total and separate I2 from multilevel-meta-analytic model
#'
#' @param model 
#' @param method 
#'
#' @return
#' @export
#'
#' @examples
I2 <- function(model, method = c("Wolfgang", "Shinichi")) {
    
    ## evaluate choices
    method <- match.arg(method)
    
    # Wolfgang's method
    if (method == "Wolfgang") {
        W <- solve(model$V)
        X <- model.matrix(model)
        P <- W - W %*% X %*% solve(t(X) %*% W %*% X) %*% t(X) %*% W
        I2_total <- sum(model$sigma2)/(sum(model$sigma2) + (model$k - model$p)/sum(diag(P)))
        I2_each <- model$sigma2/(sum(model$sigma2) + (model$k - model$p)/sum(diag(P)))
        names(I2_each) = paste0("I2_", model$s.names)
        
        # putting all together
        I2s <- c(I2_total = I2_total, I2_each)
        
        # or my way
    } else {
        # sigma2_v = typical sampling error variance
        sigma2_v <- sum(1/model$vi) * (model$k - 1)/(sum(1/model$vi)^2 - sum((1/model$vi)^2))
        I2_total <- sum(model$sigma2)/(sum(model$sigma2) + sigma2_v)  #s^2_t = total variance
        I2_each <- model$sigma2/(sum(model$sigma2) + sigma2_v)
        names(I2_each) = paste0("I2_", model$s.names)
        
        # putting all together
        I2s <- c(I2_total = I2_total, I2_each)
    }
    return(I2s)
}


#' Title: the function to get prediction intervals (crediblity intervals) from rma objects (metafor)
#'
#' @param model: rma.mv object 
#' @param mod: the name of a moderator 
get_pred <- function(model, mod = " ") {
    name <- as.factor(str_replace(row.names(model$beta), mod, ""))
    len <- length(name)
    
    if (len != 1) {
        newdata <- matrix(NA, ncol = len, nrow = len)
        for (i in 1:len) {
            # getting the position of unique case from X (design matrix)
            pos <- which(model$X[, i] == 1)[[1]]
            newdata[, i] <- model$X[pos, ]
        }
        pred <- predict.rma(model, newmods = newdata)
    } else {
        pred <- predict.rma(model)
    }
    lowerPR <- pred$cr.lb
    upperPR <- pred$cr.ub
    
    table <- tibble(name = name, lowerPR = lowerPR, upperPR = upperPR)
}


```


## Load data   
```{r load data}

### Phenotypic and metadata ###

# Load effect sizes
mean.dif <- read.csv("../data/variation.ES.general.csv", head = TRUE)%>%
  # indicate dataset including TS
  mutate(data.type = "withTS")

# Observations with TS
TS <- read.csv("../data/variation.ES.TS.csv", head = TRUE) %>%
  filter(yi < 0) %>%
  distinct(ES.ID)
# Effect sizes without observations with TS
noTS <- read.csv("../data/variation.ES.general.csv", head = TRUE) %>%
  filter(!ES.ID %in% unique(TS$ES.ID)) %>%
  # indicate dataset without TS
  mutate(data.type = "withoutTS")

# combine dataset with/without TS
with.without.TS <- bind_rows(mean.dif, noTS) %>%
  mutate(metaunit =  str_c(data.type, cross, sep = "_")) %>%
  # Use observations with both reciprocal crosses
  drop_na(contains("Mn"), contains("SD"))


### Phylogeny ###

# compute branch lengths of tree
phylo_branch <- read.tree(file = "../data/phylo.variation.general.tre") %>%
  compute.brlen(bin.tree, method = "Grafen", power = 1)

# check tree is ultrametric
is.ultrametric(phylo_branch) # TRUE

# saving phylogeneic matrix
phylo_cor <- vcv(phylo_branch, cor = T)

# generating inverse phylogenetic matrix for MCMCglmm
# Note one of the tips is called "Lates calcarifer (estimated)"
phylo_branch$node.label <- NULL
phylo_MCMC <- MCMCglmm::inverseA(phylo_branch, nodes = "ALL", scale = TRUE)$Ainv

```

# Assessing dominance in variance
Compare midparent and hybrids by comparing difference in CV from spS to midparent (spL+spS/2) and to hybrids.    
phenotypic difference from spS to midparent: *((spL+spS/2)/SS)*   
phenotypic difference from spS to hybrids: *(hyb/SS)*  
As the variance of *((spL+spS/2)/SS)*, we use the variance of *(LL/SS)* for simplicity

```{r assessing dominance}

# Calculate midparent 
midparent <- with.without.TS %>%
  # Divide mean by 2 in spL whereas hybrids are constant
   mutate(
     CV.spL = SD.spL/Mn.spL,
     CV.spS = SD.spS/Mn.spS
     ) %>%
   mutate(
     lnCVR.midparent.es =
       ifelse(
         cross == "spL",
         # log(CVe/CVc)
         log((CV.spL + CV.spS)/2) - log(CV.spS) +  
           # 1/2(Ne-1) - 1/2(Nc-1)
           1/(N.spL + N.spS -2) - 1/(2*N.spS - 2), 
         yi
         )
     # Divide variance by 4 in spL whereas hybrids are constant
     ) %>%
   mutate_at("cross", as.factor) %>%
   within(levels(cross) <- c("hybLS", "hybSL", "midparent"))

################################################
# phylogenetic random regresson (ANOVA)
################################################

for (j in c("withTS", "withoutTS")) {

  # Regress by lnRR of phenotypic difference between parentals 
  midparent.random <- rma.mv(
    yi = yi, V = vi, 
    data = midparent %>% 
      filter(data.type == j), 
    method = "REML",
    random = list(~1 | spL.name, ~1 | Study.ID, ~1 | Cross.ID),
    R = list(spL.name = phylo_cor),
    mods = ~ relevel(cross, ref = "midparent")
    )
  
  print(j)    
  print(summary(midparent.random))

  print(paste("hybLS is larger than midparent in",
      round(
        (exp(midparent.random$beta[1]+midparent.random$beta[2]) -
           exp(midparent.random$beta[1])
         )*100,
        2),
      "percent"))
  print(paste("hybSL is larger than midparent in",
      round(
        (exp(midparent.random$beta[1]+midparent.random$beta[3]) -
           exp(midparent.random$beta[1])
         )*100,
        2),
      "percent"))
  
  
  #############################################
  # Meta-analysis for midparent to plot band
  #############################################
  
  # Phylogenetic random meta-analysis
  midparent.random <- rma.mv(
    yi = lnCVR.midparent.es, 
    V = vi, 
    data = midparent %>%
      filter(cross == "midparent" | data.type == j), 
    method = "REML",
    random = list(~1 | spL.name, ~1 | Study.ID, ~1 | Cross.ID, ~1 | ES.ID),
    R = list(spL.name = phylo_cor)
    )
  
  # creating summary table of results
  pred_ma <- get_pred(midparent.random)
  assign(
    paste("midparent.ma", j, sep = "."), 
    cbind(
      midparent.random$beta,   # regression: estimate
      midparent.random$ci.lb,  # regression: 95ci 
      midparent.random$ci.ub,  # regression: 95ci
      pred_ma$lowerPR,    # lower prediction range
      pred_ma$upperPR    # lower prediction range
      ) %>%
      as.data.frame() %>%
      rename(
        estimate = V1, lowerCI = V2, upperCI = V3, lowerPR = V4, upperPR = V5
        ) %>%
      mutate(data.type = j)
  )
}

### Midparent meta-analysis result for plot below ###
midparent.ma <- bind_rows(
  midparent.ma.withTS, midparent.ma.withoutTS
)

```

Both reciprocal hybrids had smaller variance than midparent-of-variance

# Meta-analysis for each hybrids and parental
Calculate meta-analytic mean of lnCVR between less varied parental and the other crosses (hybrids and more varied parental)

```{r meta-analysis for each}

for (j in c("withTS", "withoutTS")) {
  for (h in c("spL", "hybLS", "hybSL")) {
  
    # Make data for each cross
    split.data <- with.without.TS %>%
      # use observations which both crosses are availabe to compare the heterogeneity of both crosses
      filter(
        cross == h,
        data.type == j
        )
    
    print(paste(h))
    
    # Phylogenetic random meta-analysis
    phyl.random <- rma.mv(
      yi = yi, 
      V = vi, 
      data = split.data, method = "REML",
      random = list(~1 | spL.name, ~1 | Study.ID, ~1 | Cross.ID, ~1 | ES.ID),
      R = list(spL.name = phylo_cor)
      )
    
    # Funnel plot
    funnel(phyl.random)
    
    # creating summary table of results
    pred_ma <- get_pred(phyl.random)
    assign(
      paste("res", h, sep = "."), 
      cbind(
        phyl.random$beta,   # regression: estimate
        phyl.random$ci.lb,  # regression: 95ci 
        phyl.random$ci.ub,  # regression: 95ci
        phyl.random$pval,   # P
        length(split.data$ES.ID), # N
        pred_ma$lowerPR,    # lower prediction range
        pred_ma$upperPR    # lower prediction range
        ) %>%
        as.data.frame() %>%
        rename(
          estimate = V1, lowerCI = V2, upperCI = V3, P = V4, 
          N = V5, lowerPR = V6, upperPR = V7) %>%
        mutate(
          cross = h,    # indicate cross name (spL, hybLS, hybSL)
          data.type = j # indicate dataset without TS
          ) %>%
        mutate(metaunit =  str_c(data.type, cross, sep = "_"))
        )
    
    print(paste(
      "I^2 =", 
      I2(phyl.random)["I2_total"] %>%
        round(3)
        ))

  }
  
  assign(
    paste("res.ma", j, sep = "."),
    res.ma <- bind_rows(
      res.spL, res.hybLS, res.hybSL
      ) %>%
    as.data.frame()
  )
  
  ### Midparent estimate & CI ###
  midparent.ma.2 <- midparent.ma %>%
    filter(data.type == j)
  
  # creating a forest plot
  plot.metamean <- ggplot(
    data = res.ma, 
    aes(x = tanh(estimate), y = cross)
    ) +
    scale_x_continuous(
      limits = c(-1, 1), 
      breaks = seq(-1, 1, by = 0.2)
      ) +
    ### Midparent CI ###
    geom_rect(
      aes(
        xmin = midparent.ma.2$lowerCI,
        xmax = midparent.ma.2$upperCI,
        ymin = -Inf, ymax = Inf
        ),
      fill="Grey 95", inherit.aes = FALSE
      ) +
    ### Midparent estimate ###
    geom_vline(
      xintercept = midparent.ma.2$estimate, colour = "grey20", 
      alpha = 0.3, size = 1
      ) +
    ### Orchard plot ###
    geom_quasirandom(
      data = with.without.TS %>%
        filter(data.type == j), 
      aes(x = tanh(yi), y = cross, size = 1/vi, color = cross),
      alpha = 0.2, groupOnX = FALSE
      ) + 
    scale_color_manual(values = c("#DC267F", "#785EF0", "#FE6100")) +
    # Delete legend for colors
    guides(fill = "none", colour = "none") +
    ### Confidential interval ###
    geom_errorbarh(
      aes(xmin = tanh(lowerCI), xmax = tanh(upperCI)), 
      height = 0, size = 1.2, alpha = 0.6
      ) + 
    ### Prediction interval ### 
    geom_errorbarh(
      aes(xmin = tanh(lowerPR), xmax = tanh(upperPR)), 
      height = 0, size = 0.5, alpha = 0.6
      ) +
    geom_vline(xintercept = 0, linetype = 2, colour = "black", alpha = 0.3) + 
    ### Estimate ###
    geom_point(size = 3, shape = 21, fill = "black") + 
    ### N ###
    annotate(
      "text", x = -0.93, y = c(1.15, 2.15, 3.15),
      label = paste("italic(N)==", res.ma$N), 
      parse = TRUE, hjust = "left", size = 3.5
      ) + 
    ylab("") +
    theme_bw() +
    theme(
      legend.position = c(0, 1),
      legend.justification = c(0, 1),
      legend.direction = "horizontal",
      legend.background = element_blank()
      )

  print(plot.metamean)
  
  ggsave(
    plot = plot.metamean,
    file = paste("../Analysis/variation.general.metamean", j, "svg", sep = "."), 
  height = 5, width = 6
  )

}

# res.all.ma <- bind_rows(res.ma.withTS, res.ma.withoutTS) %>%
#   as.data.frame()

```


# comparing mean across all crosses
Comparing the differences between parentals and hybrids

```{r comparing mean across all crosses}

for (j in c("withTS", "withoutTS")) {

  mean.dif <- with.without.TS %>% 
    filter(data.type == j)


################################################
# phylogenetic random regresson (ANOVA)
################################################
# Regress by lnRR of phenotypic difference between parentals 
phyl.random.spL <- rma.mv(
  yi = yi, V = vi, 
  data = mean.dif, method = "REML",
  random = list(~1 | spL.name, ~1 | Study.ID, ~1 | Cross.ID),
  R = list(spL.name = phylo_cor),
  mods = ~ relevel(cross, ref = "spL")
  )

phyl.random.hybLS <- rma.mv(
  yi = yi, V = vi, 
  data = mean.dif, method = "REML",
  random = list(~1 | spL.name, ~1 | Study.ID, ~1 | Cross.ID),
  R = list(spL.name = phylo_cor),
  mods = ~ cross
  )


summary(phyl.random.spL)
summary(phyl.random.hybLS)
}

```



# Meta-analysis & regression for lnCVR in hybrids crosses
Which hybrids hybrid more phenotypically vary? What are the affecting factor? Compare lnCVR between hybridLS-speciesS(less vary) vs hybridSL-speciesL(more vary)  

```{r maternal vs paternal}

for (j in c("withTS", "withoutTS")) {

      hybrids <- with.without.TS %>%
        filter(
          cross != "spL",
          data.type == j
          )
  
  
  phyl.random <- rma.mv(
    yi = yi, 
    V = vi, 
    data = hybrids, method = "REML",
    random = list(~1 | spL.name, ~1 | Study.ID, ~1 | Cross.ID),
    R = list(spL.name = phylo_cor),
    mods = ~ cross*hetero.sex + divergence.COI + sq.divergence.COI + pheno.div.ratio
    )
  
  # print(summary(phyl.random))
  
  ################################################
  # Output result
  ################################################
  
  reg <- cbind(
    phyl.random$beta,   # regression: estimate
    phyl.random$ci.lb,  # regression: 95ci 
    phyl.random$ci.ub,  # regression: 95ci
    phyl.random$pval   # P
    ) %>%
    cbind(., row.names(.)) %>%
    as.data.frame() %>%
    rename(estimate = V1, ci.lb = V2, ci.ub = V3, P = V4, factors = V5) %>%
    mutate_all(as.character) %>%
    mutate_at(vars(estimate:P), as.numeric) %>%
    mutate_at(vars(factors), as.factor) %>%
    mutate(significance = ifelse(P < 0.05, "*", ""))
  
  
    plot <- ggplot(reg, aes(y = estimate, x = factors)) +
      geom_errorbar(
        aes(ymin = ci.lb, ymax = ci.ub, colour=significance), 
        width=.0001
        ) +
      scale_colour_manual(values = c("black", "#D41159"), aesthetics = "colour") +
      geom_point(aes(colour=significance))+
      scale_fill_manual(values = c("black", "#D41159"), aesthetics = "colour") +
      scale_x_discrete(limits = rev(reg$factors))+ # reverse x-axis
      geom_hline(yintercept = 0, linetype="dashed", size = 0.3) +
      coord_flip() +
      ylab("") + xlab(j) +
      mytheme +
      theme(legend.position = "none")
  
  print(plot)
  
  print(
    reg %>%
      as.data.frame() %>%
      mutate_if(is.numeric, round, 3) %>%
      column_to_rownames("factors")
    )
  
  if(j == "withTS") {
    ggsave(
      plot = plot, 
      file = "../Analysis/variation.general.result.svg", 
      height = 4.5, width = 4
      )  
  } 

}

```

Significant effect was "both crosses of hybrid tend to become vary as parental species phenotypically diverges" only.
